---
title: "2023 Mid-Year Conduit Filing Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    latex_engine: lualatex
    keep_tex: true
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Arial}
  - \usepackage{xcolor}
  - \usepackage{titlesec}
  - \definecolor{blue_base}{HTML}{008dc9}
  - \definecolor{blue_strong}{HTML}{0f2d5b}
  - \definecolor{blue_weak}{HTML}{90d3fb}
  - \definecolor{black_base}{HTML}{000000}
  - \definecolor{black_weak}{HTML}{595959}
  - \definecolor{black_weakest}{HTML}{cccccc}
  - \definecolor{white_base}{HTML}{ffffff}
  - \definecolor{white_weak}{HTML}{f7f7f7}
  - \definecolor{cat_0_strong}{HTML}{cc850a}
  - \definecolor{cat_0_text}{HTML}{754d06}
  - \definecolor{cat_1_base}{HTML}{f26829} 
  - \definecolor{cat_1_strong}{HTML}{e85130}
  - \definecolor{cat_1_text}{HTML}{9a3709}
  - \definecolor{cat_2_base}{HTML}{bd53bd}
  - \definecolor{cat_2_text}{HTML}{8e2f8e}
  - \definecolor{cat_3_base}{HTML}{6363c0} 
  - \definecolor{cat_3_text}{HTML}{4b4baf}
  - \definecolor{cat_4_base}{HTML}{008dc9}
  - \definecolor{cat_4_text}{HTML}{245993}
  - \definecolor{cat_5_base}{HTML}{40bf73}
  - \definecolor{cat_5_strong}{HTML}{37a463}
  - \definecolor{cat_5_text}{HTML}{1d6339}
  - \definecolor{cat_99_base}{HTML}{cccccc}
  - \definecolor{cat_99_strong}{HTML}{878787}
  - \definecolor{cat_99_text}{HTML}{595959}
  - \definecolor{ab_midnight}{HTML}{1C2345}
  - \definecolor{ab_marigold}{HTML}{FFA300}
  - \titleformat{\title}{\color{black_base}}{}{0pt}{}
  - \titleformat{\section}{\Large\color{cat_4_text}}{}{0pt}{}
  - \titleformat{\subsection}{\large\color{cat_0_strong}}{}{0pt}{}
  - \titlespacing\section{0pt}{24pt plus 4pt minus 2pt}{6pt plus 2pt minus 2pt}
  - \titlespacing\subsection{0pt}{12pt plus 4pt minus 2pt}{2pt plus 2pt minus 2pt}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
  - \pagestyle{fancy}
  - \fancyhead[LO,LE]{Mid-Year Conduit Filing Analysis} 
  - \fancyhead[RO,RE]{Page \thepage \hspace{1pt} of \pageref{LastPage}}
  - \fancyfoot{}
  - \fancyfoot[LO,LE]{sarah.mathey@gmail.com}
  - \color{cat_99_text}

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=12, fig.height=8, fig.align = "center", warning = FALSE)

#Load libraries & source code
library(tidyverse)
library(DBI)
library(RSQLite)
library(flextable)
library(scales)
library(lubridate)
library(extrafont)
library(patchwork)

source("viz_theme.R")
source("file_paths.R")
source("sql_queries_ab_2019.R")
source("sql_queries_ab_2023.R")
source("sql_queries_wr_2023.R")
source("sql_queries_wr_ab_xcycle.R")

loadfonts()

# Initialize SQL connection
database_file <- database_path
conn <- dbConnect(RSQLite::SQLite(), dbname = database_file)

# Set flextable defaults 
set_flextable_defaults(
  font.size = 10,
  font.family = "Arial",
  theme_fun = theme_zebra,
  padding = 6)

#Pull in basic data 

ttl_raised_ab_2019_my <- ttl_raised_ab_2019_my%>%dbGetQuery(conn, .)

ttl_raised_ab_2023_my <- ttl_raised_ab_2023_my%>%dbGetQuery(conn, .)

ttl_raised_wr_2023_my <- ttl_raised_wr_2023_my%>%dbGetQuery(conn, .)

```

# Introduction 

ActBlue and WinRed act as conduits for much of the online contributions for federal political committees on the left and right, respectively. As such, they are required to file regular reports with the Federal Election Commission, detailing the contributions they received that were earmarked for other federal committees. These filings contain an incredible amount of data that can be used to better understands trends and patterns in online political giving. On July 31, 2023, both committees filed reports detailing the contributions they received for the first six months of the 2024 election cycle, from January to June 2023. 

Analyzing the data in these filings can paint a picture of the fundraising "horse race" (which committees raised more than their competitors), as well as information on who gives, when, and to whom. Comparing current filings to filings from the same periods in prior election cycles can shed light on trends and evolutions in online political giving. 

The 2020 election cycle broke records in online political fundraising. As a presidential election cycle, the 2024 cycle will be able to provide context for the 2020 fundraising numbers - was it an outlier? Will a Biden-Trump rematch lead to the same fundraising numbers produced in 2020? Which donors come back for this presidential election cycle, and which dropped off? 

While this report will not answer all of these questions, the data available in ActBlue and WinRed's filings can, and is a worthy data source for further exploration. 

# Outline

This report contains information on the following topics: 


  I. Conduit Toplines and Daily Fundraising
  
  II. Top Committees and Fundraising Days
  
  III. Presidential Fundraising


\newpage

# Conduit Toplines and Daily Fundraising

## Toplines 

```{r toplines, echo=FALSE}
ab_2019_toplines <-ab_2019_toplines %>%dbGetQuery(conn, .)
ab_2019_toplines<- ab_2019_toplines%>%mutate(avg_contribution = amt_raised/n_contributions, row_name="ActBlue 2019 MY")

ab_2023_toplines <-ab_2023_toplines %>%dbGetQuery(conn, .)
ab_2023_toplines<- ab_2023_toplines%>%mutate(avg_contribution = amt_raised/n_contributions, row_name="ActBlue 2023 MY")

wr_2023_toplines <-wr_2023_toplines %>%dbGetQuery(conn, .)
wr_2023_toplines<- wr_2023_toplines%>%mutate(avg_contribution = amt_raised/n_contributions, row_name="WinRed 2023 MY")

conduit_toplines<- bind_rows(ab_2019_toplines, ab_2023_toplines, wr_2023_toplines)

conduit_toplines_ft<-flextable(conduit_toplines,
              col_keys = c("row_name","n_contributions", "amt_raised", "avg_contribution","n_donors",  "n_committees"))%>%
  set_header_labels(row_name = "",n_contributions = "Contributions", amt_raised = "Total Raised",avg_contribution = "Avg. Contribution",n_donors = "Donors",n_committees = "Committees")%>%
  align(i = 1, j = NULL, align = "center", part = "header")%>%
  colformat_double(j = c(3),digits = 0,prefix = "$",suffix = "")%>%
  colformat_double(j = c(4),digits = 2,prefix = "$",suffix = "")%>%
  width(j= c(2,3,4,5), 1.5, unit="in")%>%
  bold(j = 1)%>%
  bg(bg = "#e2eaf1", part = "header")%>%
  add_header_lines( values = c("Conduit Fundraising Toplines January - June") ) 

conduit_toplines_ft
```
In the first six months of 2023:

- Federal contributions processed through ActBlue's conduit decreased `r percent((ab_2019_toplines$amt_raised-ab_2023_toplines$amt_raised)/(ab_2019_toplines$amt_raised),accuracy=0.1)` compared to the first six months of 2019. 
- Donors contributed `r percent((ab_2023_toplines$amt_raised-wr_2023_toplines$amt_raised)/(wr_2023_toplines$amt_raised),accuracy=0.1)` more through ActBlue's federal conduit than WinRed's
- More than twice as many donors contributed more than double the contributions to over a thousand more committees, through ActBlue's federal conduit than through WinRed's

```{r by day cumulative, eval=TRUE, warning=FALSE, include=TRUE, echo=FALSE}

ab_by_day <-ab_daily %>%dbGetQuery(conn, .)
ab_by_day<-ab_by_day%>%mutate(yday=lubridate::yday(contribution_date), running_total = cumsum(total_raised))

ab_by_day_19<-ab_by_day%>%filter(contribution_date<='2019-12-31')
ab_by_day_23<-ab_by_day%>%filter(contribution_date>'2019-12-31')
ab_by_day_23<-ab_by_day_23%>%mutate(running_total_fixed=cumsum(total_raised))

wr_by_day <-wr_daily %>%dbGetQuery(conn, .)
wr_by_day<-wr_by_day%>%mutate(yday=lubridate::yday(contribution_date), running_total = cumsum(total_raised))

ggplot()+
geom_line(data=ab_by_day_19,aes(x=yday,y=running_total, color="ab_2019"), group=1)+
geom_line(data=ab_by_day_23,aes(x=yday,y=running_total_fixed, color="ab_2023"), group=1)+
geom_line(data=wr_by_day,aes(x=yday,y=running_total, color="wr_2023"),  group=1)+
scale_x_continuous(breaks = c(1,31,59,90,120,151,181), labels= c("Jan","Feb","Mar","Apr","May","Jun","Jul") )+
scale_y_continuous(limits=c(0,400000000), breaks = c(0, 100000000,200000000,300000000,400000000), labels= c("$0M","$100M","$200M","$300M","$400M"))+
scale_color_manual(values=c(hex$blue_weak,hex$blue_base,hex$biden_red_75),labels = c("ActBlue 2019", "ActBlue 2023", "WinRed 2023"),name="")+
xlab("Month")+
ylab("Dollars Raised")+
labs(title="Cumulative Dollars Raised Through Conduits", subtitle="January through June")+
sjm_theme()
```
\newpage

- The impact of presidential campaign launches, and national political events like the first of the Trump indictments is evident in the data
```{r by day, eval=TRUE, include=TRUE, echo=FALSE}
ggplot()+
  geom_line(data=ab_by_day_19,aes(x=yday,y=total_raised, color="ab_2019"), group=1)+
  geom_line(data=ab_by_day_23,aes(x=yday,y=total_raised, color="ab_2023"), group=1)+
  geom_line(data=wr_by_day,aes(x=yday,y=total_raised, color="wr_2023"),  group=1)+
  scale_x_continuous(breaks = c(1,31,59,90,120,151,181), labels= c("Jan","Feb","Mar","Apr","May","Jun","Jul") )+
  scale_y_continuous(limits=c(0,12000000), breaks = c(0,4000000,8000000,12000000), labels= c("$0M","$4M","$8M","$12M"))+
  scale_color_manual(values=c(hex$blue_weak,hex$blue_base,hex$biden_red_75),labels = c("ActBlue 2019", "ActBlue 2023", "WinRed 2023"),name="")+
  geom_text(aes(label="Sanders 2019 Launch", x=45, y=7150000), size = 2, color = hex$blue_weak)+
  geom_text(aes(label="O'Rourke 2019 Launch", x=75,y=7850010), size = 2,color = hex$blue_weak)+
  geom_text(aes(label="Biden Launches", x=115, y=7850000), size = 2, color = hex$blue_base)+
  geom_text(aes(label="Trump Legal News", x=100, y=5300010), size = 2, color = hex$biden_red_75)+
  xlab("Month")+
  ylab("Dollars Raised")+
  labs(title="Daily Dollars Raised Through Conduits", subtitle="January - June")+
  sjm_theme()

```
\newpage

## Fundraising by Committee Type 

The differences between fundraising through ActBlue in the first six months of 2020 election cycle and the 2023 elections is driven in part by the absence of a competitive democratic presidential primary in 2023

```{r by cmte type ab, echo=FALSE, eval=TRUE, include=TRUE}
by_cmte_ab_2019_my <- by_cmte_ab_2019_my %>%dbGetQuery(conn, .)
by_cmte_ab_2023_my <- by_cmte_ab_2023_my %>%dbGetQuery(conn, .)

ab_by_cmte_tp_2019<-by_cmte_ab_2019_my%>%
  group_by(cmte_tp)%>%
  summarise(ttl_raised=sum(total_raised))

ab_by_cmte_tp_2019$cmte_tp_clean = case_when(
  ab_by_cmte_tp_2019$cmte_tp == "S" ~ "Senate",
  ab_by_cmte_tp_2019$cmte_tp == "H" ~ "House",
  ab_by_cmte_tp_2019$cmte_tp == "P" ~ "Presidential",
  ab_by_cmte_tp_2019$cmte_tp == "X" ~ "Party Committee",
  ab_by_cmte_tp_2019$cmte_tp == "Y" ~ "Party Committee",
  ab_by_cmte_tp_2019$cmte_tp == "Z" ~ "Party Committee",
  ab_by_cmte_tp_2019$cmte_tp %in% c("I","O","U")  ~ "Independent Expenditure",
  ab_by_cmte_tp_2019$cmte_tp %in% c("N","Q","V","W")  ~ "PAC",
  ab_by_cmte_tp_2019$cmte_tp %in% c("C","E")  ~ "Communication Costs",
  ab_by_cmte_tp_2019$cmte_tp == "D"  ~ "Delegate Committee"
)

# Compute percentages
ab_by_cmte_tp_2019_clean<-ab_by_cmte_tp_2019%>%
  group_by(cmte_tp_clean)%>%
  summarise(ttl_raised=sum(ttl_raised))

ab_by_cmte_tp_2019_clean$fraction = ab_by_cmte_tp_2019_clean$ttl_raised / sum(ab_by_cmte_tp_2019_clean$ttl_raised)

# Compute the cumulative percentages (top of each rectangle)
ab_by_cmte_tp_2019_clean$ymax = cumsum(ab_by_cmte_tp_2019_clean$fraction)

# Compute the bottom of each rectangle
ab_by_cmte_tp_2019_clean$ymin = c(0, head(ab_by_cmte_tp_2019_clean$ymax, n=-1))

# Compute label position
ab_by_cmte_tp_2019_clean$labelPosition <- (ab_by_cmte_tp_2019_clean$ymax + ab_by_cmte_tp_2019_clean$ymin) / 2

# Compute a good label
ab_by_cmte_tp_2019_clean$label <- paste0(ab_by_cmte_tp_2019_clean$cmte_tp_clean, "\n : ", ab_by_cmte_tp_2019_clean$fraction)

ab_by_cmte_tp_2019_clean$cycle <- "2019"

# 2023
ab_by_cmte_tp_2023<-by_cmte_ab_2023_my%>%
  group_by(cmte_tp)%>%
  summarise(ttl_raised=sum(total_raised))

ab_by_cmte_tp_2023$cmte_tp_clean = case_when(
  ab_by_cmte_tp_2023$cmte_tp == "S" ~ "Senate",
  ab_by_cmte_tp_2023$cmte_tp == "H" ~ "House",
  ab_by_cmte_tp_2023$cmte_tp == "P" ~ "Presidential",
  ab_by_cmte_tp_2023$cmte_tp == "X" ~ "Party Committee",
  ab_by_cmte_tp_2023$cmte_tp == "Y" ~ "Party Committee",
  ab_by_cmte_tp_2023$cmte_tp == "Z" ~ "Party Committee",
  ab_by_cmte_tp_2023$cmte_tp %in% c("I","O","U")  ~ "Independent Expenditure",
  ab_by_cmte_tp_2023$cmte_tp %in% c("N","Q","V","W")  ~ "PAC",
  ab_by_cmte_tp_2023$cmte_tp %in% c("C","E")  ~ "Communication Costs",
  ab_by_cmte_tp_2023$cmte_tp == "D"  ~ "Delegate Committee"
)
# Compute percentages
ab_by_cmte_tp_2023_clean<-ab_by_cmte_tp_2023%>%
  group_by(cmte_tp_clean)%>%
  summarise(ttl_raised=sum(ttl_raised))

ab_by_cmte_tp_2023_clean$fraction = ab_by_cmte_tp_2023_clean$ttl_raised / sum(ab_by_cmte_tp_2023_clean$ttl_raised)

# Compute the cumulative percentages (top of each rectangle)
ab_by_cmte_tp_2023_clean$ymax = cumsum(ab_by_cmte_tp_2023_clean$fraction)

# Compute the bottom of each rectangle
ab_by_cmte_tp_2023_clean$ymin = c(0, head(ab_by_cmte_tp_2023_clean$ymax, n=-1))

# Compute label position
ab_by_cmte_tp_2023_clean$labelPosition <- (ab_by_cmte_tp_2023_clean$ymax + ab_by_cmte_tp_2023_clean$ymin) / 2

# Compute a good label
ab_by_cmte_tp_2023_clean$label <- paste0(ab_by_cmte_tp_2023_clean$cmte_tp_clean, "\n : ", ab_by_cmte_tp_2023_clean$fraction)
 
ab_by_cmte_tp_2023_clean$cycle <- "2023"

# Make the plot
ab_by_cmte_tp_xcycle<-bind_rows(ab_by_cmte_tp_2019_clean,ab_by_cmte_tp_2023_clean)


ab_by_cmte_tp_xcycle %>%
ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=cmte_tp_clean)) +
  facet_wrap(~cycle) +
  geom_rect() +
  coord_polar(theta="y") + 
  xlim(c(2, 4))+
  geom_text( x=3.5, aes(y=labelPosition, label=paste(round(fraction*100.00,1),'%')), color=hex$white_weak, size=3) +
  sjm_theme()+
  theme_void()+
  scale_fill_manual(values=c(hex$blue_weak,hex$cat_2_base,hex$cat_3_base,hex$cat_4_text,hex$blue_strong,hex$blue_base,hex$cat_99_base))+
  theme(text           = element_text(color = hex$cat_99_text),
    legend.position    = "top",
    plot.title         = element_text(face = "bold", hjust = 0.5),
    plot.subtitle      = element_text(size = 9, margin = margin(9,0,9,0), hjust = 0.5))+
  labs(fill = "Committee Type", title="Proportion of Dollars Raised through ActBlue by Committee Type")

```
\newpage 

```{r by cmte type wr 2023, echo=FALSE, eval=TRUE, include=TRUE}
by_cmte_wr_2023_my <- by_cmte_wr_2023_my %>%dbGetQuery(conn, .)

wr_by_cmte_tp_2023<-
  by_cmte_wr_2023_my%>%group_by(cmte_tp)%>%summarise(ttl_raised=sum(total_raised))

wr_by_cmte_tp_2023$cmte_tp_clean = case_when(
  wr_by_cmte_tp_2023$cmte_tp == "S" ~ "Senate",
  wr_by_cmte_tp_2023$cmte_tp == "H" ~ "House",
  wr_by_cmte_tp_2023$cmte_tp == "P" ~ "Presidential",
  wr_by_cmte_tp_2023$cmte_tp == "X" ~ "Party Committee",
  wr_by_cmte_tp_2023$cmte_tp == "Y" ~ "Party Committee",
  wr_by_cmte_tp_2023$cmte_tp == "Z" ~ "Party Committee",
  wr_by_cmte_tp_2023$cmte_tp %in% c("I","O","U")  ~ "Independent Expenditure",
  wr_by_cmte_tp_2023$cmte_tp %in% c("N","Q","V","W")  ~ "PAC",
  wr_by_cmte_tp_2023$cmte_tp %in% c("C","E")  ~ "Communication Costs",
  wr_by_cmte_tp_2023$cmte_tp == "D"  ~ "Delegate Committee"
)

# Compute percentages
wr_by_cmte_tp_2023_clean<-wr_by_cmte_tp_2023%>%group_by(cmte_tp_clean)%>%summarise(ttl_raised=sum(ttl_raised))

wr_by_cmte_tp_2023_clean$fraction = wr_by_cmte_tp_2023_clean$ttl_raised / sum(wr_by_cmte_tp_2023_clean$ttl_raised)

# Compute the cumulative percentages (top of each rectangle)
wr_by_cmte_tp_2023_clean$ymax = cumsum(wr_by_cmte_tp_2023_clean$fraction)

# Compute the bottom of each rectangle
wr_by_cmte_tp_2023_clean$ymin = c(0, head(wr_by_cmte_tp_2023_clean$ymax, n=-1))

# Compute label position
wr_by_cmte_tp_2023_clean$labelPosition <- (wr_by_cmte_tp_2023_clean$ymax + wr_by_cmte_tp_2023_clean$ymin) / 2

# Compute a good label
wr_by_cmte_tp_2023_clean$label <- paste0(wr_by_cmte_tp_2023_clean$cmte_tp_clean, "\n : ", wr_by_cmte_tp_2023_clean$fraction)
 
# Make the plot
wr_by_cmte_tp_2023_clean %>%
ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=cmte_tp_clean)) +
  geom_rect() +
  coord_polar(theta="y") + 
  xlim(c(2, 4)) +
  geom_text( x=3.5, aes(y=labelPosition, label=paste(round(fraction*100.00,1),'%')), color=hex$white_weak, size=3) +
  sjm_theme()+
  theme_void()+
  scale_fill_manual(values=c("#660d04","#a91608","#ed1e0c","#f5594b","#f7786d","#ffafaf",hex$cat_99_base))+
  theme(text           = element_text(color = hex$cat_99_text),
    legend.position    = "top",
    plot.title         = element_text(face = "bold", hjust = 0.5),
    plot.subtitle      = element_text(size = 9, margin = margin(9,0,9,0), hjust = 0.5))+
  labs(fill = "Committee Type", title="Proportion of Dollars Raised through 2023 Winred by Committee Type")

```
- WinRed's 2023 mid year filing reflects Trump's use of a joint fundraising PAC as his main fundraising arm. The remaining Presidential candidates don't even match house fundraising through WinRed, in contrast to the impact of the Democratic Presidential Primary fundraising seen in 2019 through ActBlue.

\newpage

# Top Committees and Fundraising Days

## Top Committees 

- In the first six months of 2023, Trump and the Republican National Committees represented 44.3% of WinRed's total fundraising. Comparatively, Biden and the Democratic National Party Committees represented 15.2% of ActBlue's total fundraising.
- Adam Schiff's California Senate campaign was the top fundraiser on ActBlue in the first six months of 2023.


```{r top conduit committees 2023, echo=FALSE}
top_ab_cmte_23<-by_cmte_ab_2023_my%>%
  filter(cmte_nm !="Contribution to ActBlue")%>%
  arrange(desc(total_raised))%>%
  slice(1:5)%>%
  mutate(conduit="AB")

top_wr_cmte_23<-by_cmte_wr_2023_my%>%
  filter(cmte_nm !="Contribution to ActBlue")%>%
  arrange(desc(total_raised))%>%
  slice(1:5)%>%
  mutate(conduit="WR")

top_conduit_cmte<-bind_rows(top_ab_cmte_23,top_wr_cmte_23)

top_conduit_cmte%>%
  ggplot()+
  geom_col(aes(total_raised,reorder(cmte_nm, total_raised), fill=conduit), group = 1,  width = .75)+
  geom_text(aes(0, y = cmte_nm, label = dollar(round(total_raised/1000000,1),suffix="M")),hjust = 0,nudge_x = 0.3,colour = "white",size = 3)+
  scale_x_continuous(limits = c(0,50000000), breaks = c(0,10000000,20000000,30000000,40000000,50000000), labels = c("$0M", "$10M", "$20M","$30M","$40M",  "$50M"))+
  scale_fill_manual(values = c(hex$biden_blue,hex$biden_red))+
  scale_y_discrete(labels = label_wrap(20)) +
  sjm_theme()+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none")  +
  labs(title="Top 5 Committees for Each Conduit", subtitle = "Janaury - June 2023")
```
## Top Fundraising Days

```{r conduit top days, echo=FALSE, eval=TRUE, include=TRUE}
ab_top_days_23<-ab_by_day_23%>%arrange(desc(total_raised))%>%slice(1:5)%>%mutate(conduit="AB")

#conduit_top_days<-bind_rows(ab_top_days_23,wr_top_days_23)
ab_plt<-ab_top_days_23%>%
  ggplot()+
  geom_col(aes(total_raised,reorder(contribution_date, total_raised)), fill=hex$biden_blue, width = .75)+
   geom_text(aes(0, y = contribution_date, label = dollar(round(total_raised/1000000,1),suffix="M")), hjust = 0, nudge_x = 0.3, colour = "white", size = 3)+
  scale_x_continuous(limits = c(0,10000000), breaks = c(0,5000000,10000000), labels = c("$0M","$5M", "$10M"))+
  scale_y_discrete(labels = label_wrap(20)) +
  sjm_theme()+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none")  +
  ggtitle("Top 5 Days for ActBlue", subtitle = "Janaury - June 2023")

wr_top_days_23<-wr_by_day%>%
  arrange(desc(total_raised))%>%
  slice(1:5)%>%
  mutate(conduit="WR")
  
wr_plt<-wr_top_days_23%>%
  ggplot()+
  geom_col(aes(total_raised,reorder(contribution_date, total_raised)), fill=hex$biden_red, width = .75)+
  geom_text(aes(0, y = contribution_date, label = dollar(round(total_raised/1000000,1),suffix="M")), hjust = 0, nudge_x = 0.3, colour = "white", size = 3)+
  scale_x_continuous(limits = c(0,10000000), breaks = c(0,5000000,10000000), labels = c("$0M","$5M", "$10M"))+
  scale_y_discrete(labels = label_wrap(20)) +
  sjm_theme()+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none") +
  ggtitle("Top 5 Days for WinRed", subtitle = "Janaury - June 2023")

ab_plt + wr_plt

```
\newpage 

# Presidential Fundraising

```{r pres_launches_01, echo=FALSE}
pres_launches <- ab_pres_launches %>%dbGetQuery(conn, .)
```

## Biden 2023 Campaign Launch

President Biden launched his reelection campaign on April 25, 2023, exactly four years after first announcing his 2020 candidacy. 

On April 25, 2023 Biden for President raised `r dollar(as.numeric(pres_launches[pres_launches['contribution_date']=="2023-04-25"][6]))` through ActBlue. `r percent((as.numeric(pres_launches[pres_launches['contribution_date']=="2019-04-25"][6])-as.numeric(pres_launches[pres_launches['contribution_date']=="2023-04-25"][6]))/as.numeric(pres_launches[pres_launches['contribution_date']=="2019-04-25"][6]), accuracy = .1)` less than his 2019 ActBlue fundraising total of `r dollar(as.numeric(pres_launches[pres_launches['contribution_date']=="2019-04-25"][6]))`. Through June 30, 2023, the Biden for President campaign launch was the team's largest fundraising day through ActBlue. 

Through June 30, 2023, Biden for President reported $8.6M in total dollars raised from individuals `r percent(by_cmte_ab_2023_my%>%filter(recipient_cmte_id=="C00703975")%>%select(total_raised)%>%pull()/(15229734.75-6593709.54), accuracy = 0.1)` of this was raised through ActBlue. Of all the money raised from individuals, `r percent((9056016-3695717.15)/(15229734.75-6593709.54), accuracy=0.1)` came from unitemized donations. 

ActBlue's top 5 Presidential launch days from the 2020 cycle are listed in the table below. 

```{r pres_launches, echo=FALSE}

pres_launch_ft<-pres_launches%>%
  arrange(desc(total_raised))%>%
  slice(1:5)%>%
  mutate(cmte_nm = str_to_title(cmte_nm))%>%
  flextable(col_keys = c("cmte_nm",  "contribution_date", "total_raised" ))%>%
  set_header_labels(cmte_nm = "Committee Name", contribution_date = "Launch Date", total_raised = "Total Raised")%>%
  colformat_double(j = 3, digits = 0,prefix = "$", suffix = "")%>%
  align(i = 1,  j = NULL, align = "center", part = "header")%>%
  bg(bg = "#e2eaf1", part = "header")%>%
  width(j=c(1), 2, unit = 'in')%>%
  width(j=c(2), 1.5,unit = 'in')

pres_launch_ft

```

- *Note: People Powered Action was formerly known as Beto for America*

## Presidential Front Runners 

Due to the fact that President Trump is primarily fundraising through a PAC, the plots below contain data from both President Biden and President Trump's principal campaign committees, Biden for America and Donald J. Trump for President, as well as their joint fundraising committees, Biden Victory and Biden Action, and Trump Make America Great Again Committee, Trump Victory, and Trump Save America Joint Fundraising Committee. 

```{r Pres Daily bar, echo=FALSE}
biden_daily<-biden_my%>%dbGetQuery(conn, .)
trump_daily<-trump_my%>%dbGetQuery(conn, .)

biden_daily_sum<-biden_daily%>%
  group_by(contribution_date)%>%
  summarise(total_grp_raised=sum(total_raised))

trump_daily_sum<-trump_daily%>%
  group_by(contribution_date)%>%
  summarise(total_grp_raised=sum(total_raised))

pres_daily_sum<-trump_daily_sum%>%
  left_join(biden_daily_sum,by="contribution_date",suffix=c("_trump","_biden"))

pres_daily_sum<-pres_daily_sum%>%
  pivot_longer(!contribution_date, names_to = "candidate", values_to = "total_raised")

pres_daily_sum%>%
  filter(contribution_date>="2023-04-25")%>%
  ggplot() +
  geom_col(aes(contribution_date,total_raised,fill=candidate),position="dodge") +
  scale_x_discrete(breaks = c("2023-01-01","2023-02-01","2023-03-01","2023-04-01","2023-05-01","2023-06-01"), labels= c("Jan","Feb","Mar","Apr","May","Jun")) +
  scale_y_continuous(limits=c(0,3000000), breaks = c(0,1000000,2000000,3000000), labels= c("$0M","$1M","$2M","$3M")) +
  scale_fill_manual(values = c(hex$biden_blue,hex$biden_red), labels ="", name="") +
  ylab("Dollars Raised") +
  xlab("Month") +
  ggtitle("Daily Dollars Raised by Trump and Biden and their JFCs", subtitle = "March 25 - June 30, 2023") +
  sjm_theme() +
  theme(legend.position="none")

```
```{r Pres Daily cmte breakout, echo=FALSE}
biden_2023<-biden_daily%>%filter(contribution_date>="2023-01-01")

ggplot()+
    # Trump PCC
    geom_line(data=filter(trump_daily,contribution_date>="2023-04-25"& recipient_cmte_id=="C00828541"), aes(x=contribution_date, y=total_raised), group=1, color=hex$biden_red, size=0.75)+
  # Trump TMAGAC
    geom_line(data=filter(trump_daily,contribution_date>="2023-04-25"& recipient_cmte_id=="C00618371"), aes(x=contribution_date, y=total_raised), group=1, color=hex$biden_red_75, size=0.75)+
  # Trump Victory
    geom_line(data=filter(trump_daily,contribution_date>="2023-04-25"& recipient_cmte_id=="C00618389"), aes(x=contribution_date, y=total_raised), group=1, color=hex$biden_red_25, size=0.75)+
  # Trump Save America
    geom_line(data=filter(trump_daily,contribution_date>="2023-04-25"& recipient_cmte_id=="C00770941"), aes(x=contribution_date, y=total_raised), group=1, color="#ffafaf", size=0.75)+
  # Biden PCC
  geom_line(data=filter(biden_2023,contribution_date>="2023-04-25" & recipient_cmte_id=="C00703975"), aes(x=contribution_date, y=total_raised), group=1, color=hex$biden_blue, size=0.75)+
  # Biden Victory
  geom_line(data=filter(biden_2023,contribution_date>="2023-04-25" & recipient_cmte_id=="C00744946"), aes(x=contribution_date, y=total_raised), group=1, color=hex$biden_blue_75, size=0.75)+
  # Biden Action
  geom_line(data=filter(biden_2023,contribution_date>="2023-04-25" & recipient_cmte_id=="C00838912"), aes(x=contribution_date, y=total_raised), group=1, color=hex$biden_blue_25, size=0.75)+
  scale_x_discrete(breaks = c("2023-01-01","2023-02-01","2023-03-01","2023-04-01","2023-05-01","2023-06-01"), labels= c("Jan","Feb","Mar","Apr","May","Jun") )+
scale_y_continuous(limits=c(0,3000000), breaks = c(0,1000000,2000000,3000000), labels= c("$0M","$1M","$2M","$3M"))+
  geom_text(aes(x="2023-05-04",y=2900000), color=hex$biden_red, label="Donald J. Trump for President", size =3, hjust="left")+
  geom_text(aes(x="2023-05-04",y=2800000), color=hex$biden_red_75, label="Trump Make America Great Again Committee", size =3, hjust="left")+
  #geom_text(aes(x="2023-05-04",y=2600000), color=hex$biden_red_25, label="Trump Victory", size =3, hjust="left")+
  #geom_text(aes(x="2023-05-04",y=2500000), color="#ffafaf", label="Trump Save America", size =3, hjust="left")+
  geom_text(aes(x="2023-05-04",y=2700000), color=hex$biden_blue, label="Biden for America", size =3, hjust="left")+
  geom_text(aes(x="2023-05-04",y=2600000), color=hex$biden_blue_75, label="Biden Victory", size =3, hjust="left")+
  #geom_text(aes(x="2023-05-04",y=2200000), color=hex$biden_blue_25, label="Biden Action", size =3, hjust="left")+
  ylab("Dollars Raised")+
  xlab("Month")+
  ggtitle("Daily Dollars Raised by Trump and Biden and their Joint Fundraising Committees", subtitle = "March 25 - June 30, 2023")+
  sjm_theme()

```
\newpage

Beyond the total dollars raised and daily fundraising patterns, what can this data tell us about who is supporting each candidate? Political contributors must provide their occupation and employer when they donate to a federal committee. Looking at the two Presidential front runners, their are clear differences in their supporter communities, based off their reported occupations. 

``` {r, presidential occupations, echo=FALSE}
biden_donor_occ<-biden_donor_occ_query%>%dbGetQuery(conn, .)
trump_donor_occ<-trump_donor_occ_query%>%dbGetQuery(conn, .)

biden_donor_occ_plt<-biden_donor_occ%>%
  filter(occupation!="NOT EMPLOYED")%>%
  arrange(desc(n_donors))%>%
  slice(1:5)%>%
  ggplot() +
  geom_col(aes(n_donors,reorder(occupation, n_donors)), fill=hex$biden_blue, width = .75) +
  geom_text(aes(0, y = occupation, label = n_donors), hjust = 0, nudge_x = 0.3, colour = "white", size = 3) +
  scale_x_continuous(limits = c(0,6000), breaks = c(0,2000,4000,6000), labels = c("$0M", "$2K","$4K","$6K")) +
  scale_y_discrete(labels = label_wrap(20)) +
  sjm_theme() +
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none")  +
  labs(caption="Excludes ActBlue default occupation: Not Employed") +
  ggtitle("Top 5 Biden Donor Occupations", subtitle = "Janaury - June 2023")


trump_donor_occ_plt<-trump_donor_occ%>%
  filter(occupation!="RETIRED")%>%
  arrange(desc(n_donors))%>%
  slice(1:5)%>%
  ggplot()+
  geom_col(aes(n_donors,reorder(occupation, n_donors)), fill=hex$biden_red, width = .75)+
  geom_text(aes(0, y = occupation, label = n_donors), hjust = 0, nudge_x = 0.3, colour = "white", size = 3)+
  scale_x_continuous(limits = c(0,6000), breaks = c(0,2000,4000,6000), labels = c("$0M", "$2K","$4K","$6K"))+
  scale_y_discrete(labels = label_wrap(20)) +
  sjm_theme()+
  theme(panel.grid.major.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position="none")  +
  labs(caption = "Excludes WinRed default occupation: Retired")+
  ggtitle("Top 5 Trump Donor Occupations", subtitle = "Janaury - June 2023")

biden_donor_occ_plt + trump_donor_occ_plt

```

\newpage 

## WinRed Presidential candidates 

The plot below tells the horse race fundraising story of the GOP presidential primary through WinRed. 
```{r JV Presidential Race, echo=FALSE}
jv_gop_pres<-by_cmte_wr_2023_my%>%
  filter(cmte_tp=="P")

jv_gop_pres<-jv_gop_pres%>%
  mutate(cmte_nm_fct =fct_reorder(cmte_nm, total_raised))

jv_gop_pres%>%
  filter(total_raised>=50000)%>%
  ggplot() +
  geom_col(aes(reorder(cmte_nm_fct, total_raised, sum),total_raised), fill=hex$biden_red_75) +
  scale_y_continuous(limits = c(0, 4000000), breaks = c(0, 1000000,2000000,3000000,4000000), labels = c("$0M","$1M","$2M","$3M","$4M")) +
  ylab("Dollars Raised") +
  xlab("Committee") +
  ggtitle("GOP Presidential Candidates >$50K Raised Through WinRed", subtitle = "January - June 2023") +
  coord_flip() +
  sjm_theme()
```

```{r, echo=FALSE}
dbDisconnect(conn)
```
# Conclusion

These are just a few of the stories to be told through ActBlue and WinRed's mid year FEC filings. 

Future work could explore Senate and House fundraising, especially large races like the California Democratic Senate Primary and donor behavior therein - how many donors gave to multiple candidates, for example - or go beyond singular filings, as much data as they contain, to take advantage of the FEC's API and calculate each conduit's market share by committee and committee type as we did with President Biden. There is an incredible amount to learn from FEC filings. 

The next filing deadline for these two organizations is January 31, 2024 at which point they will release data from the second half of the 2023 calendar year. 
